---
title: "New York City Property Sale Price Prediction"
author: "Cyrus Cai"
date: "5/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
library(tinytex)
load("rda/edx.rda")
load("rda/train_set.rda")
load("rda/test_set.rda")
```

# 1 Introduction
With the assumption that historical data contains inights for the future, we attempt to predict human behavior. The research is done in various areas, such as retail, movies, polls etc. 

## 1.1 The Chanllenge 
In this project, I will build a regression model for property sale. The model will predict the sale price for an exsiting property in New York City.  

## 1.2 Dataset: Properties Sold in New York City 
I will use **nyc-rolling-sales** as my dataset. The data set is generated by New York City Department of Finance. You can find and download the daata through the link [https://www.kaggle.com/new-york-city/nyc-property-sales].  

```{r NYC Properties head,echo=FALSE}
head(edx,5)
```

## 1.3 Goal
I will find a consice and yet accurate model. The project is aimed at an RMSE that is less than 1000K. The project is also aimed to produce an easy-to-follow and reproducible report.


# 2 Analysis
Before modeling, I create two dataset: **edx** and **validation**. The **validation** is 10% of the full dataset, and is only used for final assessement of the model. It won't be used for anywhere else in the project. The **edx** data is going to be split into two subset, **train_set** and **test_set**.

## 2.1 The Simplest Model
The simplest model, {Sale Price}=b0+b1*{Gross Square Feet}+e_i


### Performance


### Pros

### Cons

## 2.2 Model 2


### Performance


### Pros

### Cons


## 2.3 Model 3


### Performance


### Pros

### Cons


# 3 Results

## Model Summary

# 4 Conclusion

## Summary of the report


## Limitations

## Future Works

# 5 Reference
