---
title: "New York City Property Sale Price Prediction"
author: "Cyrus Cai"
date: "5/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
if(!require(broom)) install.packages("broom", repos = "http://cran.us.r-project.org")
if(!require(graphics)) install.packages("graphics", repos = "http://cran.us.r-project.org")
if(!require(UsingR)) install.packages("UsingR", repos = "http://cran.us.r-project.org")
if(!require(forcats)) install.packages("forcats", repos = "http://cran.us.r-project.org")

if(!require(tidyverse)) install.packages("tidyverse", repos = "http://cran.us.r-project.org")
if(!require(tinytex)) install.packages("tinytex", repos = "http://cran.us.r-project.org")
if(!require(dplyr)) install.packages("dplyr", repos = "http://cran.us.r-project.org")

library(broom)
library(graphics)
library(UsingR)
library(forcats)
library(dplyr)
library(tidyverse)
library(tinytex)
load("rda/train_set.rda")
load("rda/test_set.rda")
```

# 1 Introduction
With the assumption that historical data contains inights for the future, we attempt to predict human behavior. The research is done in various areas, such as retail, movies, polls etc. 

## 1.1 The Challenge 
In this project, I will build a regression model for property sale. The model will predict the sale price for an exsiting property in New York City. I will only look at the properties which is between 50 thousand and 5 millions and is one family dwelling.

## 1.2 Dataset: Properties Sold in New York City 
I will use **nyc-rolling-sales** as my dataset. The data set is generated by New York City Department of Finance. You can find and download the daata through the link [https://www.kaggle.com/new-york-city/nyc-property-sales].  

```{r NYC Properties head,echo=FALSE}
head(train_set,5)
```

## 1.3 Goal
I will find a consice and yet accurate model. The project is aimed at adjusted R square of 50% and RMSE of 10K. The project is also aimed to produce an easy-to-follow and reproducible report.


# 2 Analysis
Before modeling, I create two dataset: **edx** and **validation**. The **validation** is 10% of the full dataset, and is only used for final assessement of the model. It won't be used for anywhere else in the project. The **edx** data is going to be split into two subset, **train_set** and **test_set**.

## 2.1 The Simplest Model
The simplest model, {Sale Price}=b0+b1x{Gross Square Feet}+e_i.I only consider the gross square feet in this model. 

```{r, echo=FALSE}
RMSE <- function(true_price,predicted_price){
  sqrt(mean((true_price-predicted_price)^2))
}
```

```{r}
fit_simplest<- lm(SALE.PRICE~GROSS.SQUARE.FEET,data = train_set)

tidy(fit_simplest)
summary(fit_simplest)[9]
```
```{r}
b0<-fit_simplest$coefficients[1]
b1<-fit_simplest$coefficients[2]
```

So the model 1 can be expressed as: {Sale.Price}=`r format(round(b0,0),scientific=F)`+`r round(b1,0)`*{GROSS.SQUARE.FEET}.

The model predicts the base price for a property is `r format(round(b0,0),scientific=F)` dollars and every extra square feet is `r round(b1,0)` dollars.

### Performance

```{r, echo=FALSE}


newd<-data.frame(GROSS.SQUARE.FEET=test_set$GROSS.SQUARE.FEET)
predicted_price<- predict(fit_simplest,newd,interval="prediction")
rmse_simplest_model<-RMSE(test_set$SALE.PRICE,predicted_price[,1])  
temp_p<-test_set %>%mutate(predicted_price=b0+b1*GROSS.SQUARE.FEET) %>% 
    mutate(residul=(SALE.PRICE-predicted_price)^2) %>%  group_by(Prop_ID) %>%
    summarise(r=sqrt(mean(residul)))
MED<-median(temp_p$r)

rmse_results <- data.frame(method="gross sq ft",   RMSE_in_thousand=round(rmse_simplest_model/1000,0),
                           MED_in_thousand=round(median(temp_p$r)/1000,0), 
                           k=round(rmse_simplest_model/median(temp_p$r),0),
                           Adj_r_sq=summary(fit_simplest)[9])

```
```{r}
rmse_results 
```

### Pros
This modle is the simplest linear regression model.

### Cons
The adjusted r square is only 28%. And the RMSE for the properties in Manhattan (Borough 1) is very high.

```{r,echo=FALSE}
temp_p<-test_set %>%mutate(predicted_price=b0+b1*GROSS.SQUARE.FEET) %>% 
    mutate(residul=(SALE.PRICE-predicted_price)^2)

loop.names<-names(temp_p)
i=2

temp_p%>%dplyr::select(i,26) %>%
     rename(A=1) %>%
        group_by(A) %>%
    summarise(r=sqrt(mean(residul)),n=NROW(A))%>%
        ggplot(aes(A,r,fill=n))+geom_bar(stat = "identity") +
       labs(x=loop.names[i],y="Residual")+scale_fill_gradient(low="blue", high="red")+
      geom_text(aes(label=round(r,3)),size=3,check_overlap = TRUE,position=position_dodge(width=0.9), vjust=-0.25)+theme(axis.text.x = element_text(angle = 90, hjust = 1))+
     geom_hline(yintercept=rmse_simplest_model)

```

Therefore, I need to consider borough as a factor for the next model.


## 2.2 Model 2
In the 2nd Model,I consider GROSS.SQUARE.FEET as well as whethr the property is in Manhattan.The regression expression is y=b0+b1xGSF+b2x(I(Manhattan)xGSF.

I(Manhattan) is identity funcition. It equals to 1 if the property is in Manhattan; otherwise zero.

```{r, echo=FALSE}
temp <- train_set %>% mutate(GSFandManhattan=case_when(BOROUGH=="1"~ 1,
                                                       TRUE ~ 0     )*GROSS.SQUARE.FEET) 
```


```{r}
fit_IsManhattan<- lm(SALE.PRICE~GROSS.SQUARE.FEET+GSFandManhattan,data = temp)
tidy(fit_IsManhattan)
summary(fit_IsManhattan)[9]   
```
```{r,echo=FALSE}
#save the model coefficient for the predcition calculation
b0<-fit_IsManhattan$coefficients[1]
b1<-fit_IsManhattan$coefficients[2]
b2<-fit_IsManhattan$coefficients[3]
```

So the model can be expressed as: {Sale.Price}=`r format(round(b0,0),scientific=F)`+`r round(b1,0)`* {GROSS.SQUARE.FEET} +`r round(b2,0)` * {IS.MANHATTAN} * {GROSS.SQUARE.FEET}.

The model predicts the base price for a property is `r format(round(b0,0),scientific=F)` dollars.If  the property is not in Manhattan,every extra square feet is `r round(b1,0)` dollars. If the property is in Manhattan, every extra square feet is `r round(b1+b2,0)` dollars. 

### Performance
```{r, echo=FALSE}
newd<-test_set %>% 
    mutate(GSFandManhattan=case_when(BOROUGH=="1"~ 1,
                                     TRUE ~ 0     )*GROSS.SQUARE.FEET) %>% 
    dplyr::select(GROSS.SQUARE.FEET,GSFandManhattan) 

predicted_price<- predict(fit_IsManhattan,newd,interval="prediction")
rmse_IsManhattan_model<-RMSE(test_set$SALE.PRICE,predicted_price[,1])
```

```{r, echo=FALSE}
temp_p<-test_set %>%
    mutate(GSFandManhattan=case_when(BOROUGH=="1"~ 1,
                                     TRUE ~ 0     )*GROSS.SQUARE.FEET) %>% 
    mutate(predicted_price=b0+b1*GROSS.SQUARE.FEET+b2*GSFandManhattan) %>% 
    mutate(residul=(SALE.PRICE-predicted_price)^2) %>%
    mutate(r=sqrt(residul))

MED<-median(temp_p$r)

rmse_results <-rbind(rmse_results, 
                     data.frame(method="Is Manhattan",                                RMSE_in_thousand=round(rmse_IsManhattan_model/1000,0),                                MED_in_thousand=round(median(temp_p$r)/1000,0),                                 k=round(rmse_IsManhattan_model/median(temp_p$r),0),
                                Adj_r_sq=summary(fit_IsManhattan)[9]))

```


```{r}
rmse_results
```

### Pros
The model quantify the price difference bettween Manhattan and other boroughs in New York. The RMSE reduced and the Adjsuted R Square increased.

```{r, echo=FALSE}
loop.names<-names(temp_p)
len<-ncol(temp_p)
i=2
temp_p%>%dplyr::select(i,len-1) %>% 
        rename(A=1) %>%
        group_by(A) %>%
        summarise(r=sqrt(mean(residul)),n=NROW(A))%>%
        ggplot(aes(A,r,fill=n))+geom_bar(stat = "identity") +
        labs(x=loop.names[i],y="Residual")+scale_fill_gradient(low="blue", high="red")+
        geom_text(aes(label=round(r,3)),size=3,check_overlap = TRUE,position=position_dodge(width=0.9), vjust=-0.25)+theme(axis.text.x = element_text(angle = 90, hjust = 1))+
        geom_hline(yintercept=rmse_IsManhattan_model)
    


```

The RMSE in Manhattan siginificantly reduced from 2 miliion to 1.4 milition.

### Cons

The adjusted r square is 30%. And the RMSE for the properties is high for some ZIP codes.

```{r, echo=FALSE}
loop.names<-names(temp_p)
len<-ncol(temp_p)
i=12
temp_p%>%dplyr::select(i,len-1) %>% 
        rename(A=1) %>%
        group_by(A) %>%
        summarise(r=sqrt(mean(residul)),n=NROW(A))%>%
        ggplot(aes(A,r,fill=n))+geom_bar(stat = "identity") +
        labs(x=loop.names[i],y="Residual")+scale_fill_gradient(low="blue", high="red")+
        geom_text(aes(label=round(r,3)),size=3,check_overlap = TRUE,position=position_dodge(width=0.9), vjust=-0.25)+theme(axis.text.x = element_text(angle = 90, hjust = 1))+
        geom_hline(yintercept=rmse_IsManhattan_model)
    

```

Therefore, let's zoom in and consider ZIP code in our next model.

## 2.3 Model 3
I added a column "price per square feet", which is calulated as {SALE.PRICE}/{GROSS.SQUARE.FEET}.The following boxplot shows the how the price per square feet varies by ZIP Code.

```{r, echo= FALSE}
train_set %>% mutate(price_per_square_feet=SALE.PRICE/GROSS.SQUARE.FEET) %>% 
  ggplot(aes(x=fct_reorder(ZIP.CODE,price_per_square_feet,.desc=TRUE),y=price_per_square_feet,fill=as.factor(BOROUGH)))+
  geom_boxplot(outlier.colour="black", outlier.shape=16,
               outlier.size=1, notch=FALSE)+
  theme(axis.text.x = element_text(angle = 90, hjust = 1),legend.position = "bottom")+
  labs(x="ZIP Code",y="Price Per Square Feet",fill="Borough")
```

Since the means of price per squaer feet by ZIP Code vary between 2000 and 500 roughly, I am going to pick a number M. If the price per square feet is higher than M, it is in a High Pricing Area. 

Then how to pick a proper number? I will run a loop and use the number that gives best adjusted R square as the M.

```{r, echo=FALSE}
df<-data.frame()
for (p in seq(100,1000,by=50)) {
HighPricingArea<- train_set %>% mutate(price_per_square_feet=SALE.PRICE/GROSS.SQUARE.FEET) %>% 
  group_by(ZIP.CODE) %>% 
  summarise(m=mean(price_per_square_feet),n=NROW(Prop_ID)) %>% 
  arrange(desc(m)) %>% 
  mutate(PPSF=case_when(m>p~1,
                                        TRUE ~ 0))
temp <- train_set %>% 
                   left_join(HighPricingArea,by=c("ZIP.CODE")) %>% 
  mutate(GSFandManhattan=case_when(BOROUGH=="1"~ 1,
                                   TRUE ~ 0     )*GROSS.SQUARE.FEET,
         GSFandHIGH.PRICING.AREA=PPSF*GROSS.SQUARE.FEET)
  
fit_IsManhattanAndHighPricingArea<- lm(SALE.PRICE~GROSS.SQUARE.FEET+GSFandManhattan+ GSFandHIGH.PRICING.AREA,data = temp)

tidy(fit_IsManhattanAndHighPricingArea)
summary(fit_IsManhattanAndHighPricingArea)    
df<-rbind(df,c(p=p,R=summary(fit_IsManhattanAndHighPricingArea)[9] ) )
}
```


```{r}
df[which.max(df$R.adj.r.squared),]
df %>% ggplot(aes(x=p,y=R.adj.r.squared))+geom_point()+geom_line()
```

So M=450. 

In the 3rd Model,I consider whether the property is in a High Pricing Area.The regression expression is y=b0+b1xGSF+b2x(I(Manhattan)xGSF.+b3x(I(PPSF>450)xGSF.

I(PPSF>450) equals to 1 if the price per square feet of the property is higher than 450; otherwise zero.

```{r, echo=FALSE}
HighPricingArea<- train_set %>% mutate(price_per_square_feet=SALE.PRICE/GROSS.SQUARE.FEET) %>% 
  group_by(ZIP.CODE) %>% 
  summarise(m=mean(price_per_square_feet),n=NROW(Prop_ID)) %>% 
  arrange(desc(m)) %>% 
  mutate(PPSF=case_when(m>450~1,
                        TRUE ~ 0))
temp <- train_set %>% 
  left_join(HighPricingArea,by=c("ZIP.CODE")) %>% 
  mutate(GSFandManhattan=case_when(BOROUGH=="1"~ 1,
                                   TRUE ~ 0     )*GROSS.SQUARE.FEET,
         GSFandHIGH.PRICING.AREA=PPSF*GROSS.SQUARE.FEET)




```


```{r}
fit_IsManhattanAndHighPricingArea<- lm(SALE.PRICE~GROSS.SQUARE.FEET+GSFandManhattan+ GSFandHIGH.PRICING.AREA,data = temp)

tidy(fit_IsManhattanAndHighPricingArea)
summary(fit_IsManhattanAndHighPricingArea)[9]
```
```{r,echo=FALSE}
#save the model coefficient for the predcition calculation
b0<-fit_IsManhattanAndHighPricingArea$coefficients[1]
b1<-fit_IsManhattanAndHighPricingArea$coefficients[2]
b2<-fit_IsManhattanAndHighPricingArea$coefficients[3]
b3<-fit_IsManhattanAndHighPricingArea$coefficients[4]
```

So the model can be expressed as: {Sale.Price}=`r format(round(b0,0),scientific=F)`+`r round(b1,0)`* {GROSS.SQUARE.FEET} +`r round(b2,0)` * {IS.MANHATTAN} * {GROSS.SQUARE.FEET}+`r round(b3,0)` * {IS.HIGH.PRICING.AREA} * {GROSS.SQUARE.FEET}.

The model predicts the base price for a property is `r format(round(b0,0),scientific=F)` dollars.Every extra square feet is:

If  the property is not in Manhattan and not in high pricing area, `r round(b1,0)` dollars;

If the property is in Manhattan and not in high pricing area,  `r round(b1+b2,0)` dollars; 

If the property is not in Manhattan and in high pricing area,  `r round(b1+b3,0)` dollars; 

If the property is in Manhattan and in high pricing area,  `r round(b1+b2+b3,0)` dollars. 

### Performance
```{r, echo=FALSE}
newd<-test_set %>% 
  left_join(HighPricingArea,by=c("ZIP.CODE")) %>% 
  mutate(GSFandManhattan=case_when(BOROUGH=="1"~ 1,
                                   TRUE ~ 0     )*GROSS.SQUARE.FEET,
         GSFandHIGH.PRICING.AREA=PPSF*GROSS.SQUARE.FEET) %>%  
  dplyr::select(GROSS.SQUARE.FEET,GSFandManhattan,GSFandHIGH.PRICING.AREA) 


predicted_price<- predict(fit_IsManhattanAndHighPricingArea,newd,interval="prediction")
rmse_IsManhattanAndHighPricingArea_model<-RMSE(test_set$SALE.PRICE,predicted_price[,1])  

```
```{r,echo=FALSE}
temp_p<-test_set %>%
  left_join(HighPricingArea,by=c("ZIP.CODE")) %>% 
  mutate(GSFandManhattan=case_when(BOROUGH=="1"~ 1,
                                   TRUE ~ 0     )*GROSS.SQUARE.FEET,
         GSFandHIGH.PRICING.AREA=PPSF*GROSS.SQUARE.FEET) %>%  
  mutate(predicted_price=b0+b1*GROSS.SQUARE.FEET+b2*GSFandManhattan+b3*GSFandHIGH.PRICING.AREA) %>% 
  mutate(residul=(SALE.PRICE-predicted_price)^2) %>%
  mutate(r=sqrt(residul))

MED<-median(temp_p$r)

#Store the results 
rmse_results <-rbind(rmse_results, 
                     data.frame(method="Is Manhattan and High Pricing Area",
                                RMSE_in_thousand=round(rmse_IsManhattanAndHighPricingArea_model/1000,0),
                                MED_in_thousand=round(median(temp_p$r)/1000,0), 
                                k=round(rmse_IsManhattanAndHighPricingArea_model/median(temp_p$r),0),
                                Adj_r_sq=summary(fit_IsManhattanAndHighPricingArea)[9]))



```
```{r}
rmse_results
```


### Pros

### Cons


# 3 Results

## Model Summary

# 4 Conclusion

## Summary of the report


## Limitations

## Future Works

# 5 Reference
